\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation for Machine Learning

### 2018-05-09 ccs/eval.tex
para

\subsection{Secure prediction}
\paragraph{Standard classifiers.}
%\label{sec:shallow}
We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
decision trees, from~\cite{shafindss} on the following data sets from the UCI
machine learning repository~\cite{uci}:
 the Wisconsin Breast Cancer data set, 
Credit Approval data set, Audiology (Standardized) data set, Nursery
data set, and ECG (electrocardiogram) classification data
from~\cite{barni}.

@@

\paragraph{Deep neural nets.}
We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
comparison, we consider their implementations from
\minion~\cite{minionn}, which outperforms their previous
implementations. Table~\ref{tab:nn} shows the
results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
We note that for each of these DNNs, \minion provides a
specialized protocol, while \tool uses a generic 2PC protocol
(auto) generated from high-level code.

@@

\paragraph{State-of-the-art classifiers.}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
 $W$ is a $10\times 784$ matrix, and  $b$ is a $10$ length
vector.
We implement this classifier in \tool and present the results in the
first row of Table~\ref{tab:tf}. 
%We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines and cannot be evaluated using garbled circuits.

@@

\paragraph{Scalability.}  To
illustrate the scalability of partitioning, we evaluate a
sequence of DNNs with and without partitioning
(where every layer of each DNN is a separate stage) in~\figureref{partitioning}.
For DNNs with up to 4 (hidden) layers, the performance is almost identical and the lines overlap, thereby
illustrating that partitioning did not cause any noticeable performance overheads.
Memory issues start showing up in the non-partitioned implementation with the 5 layer DNN and it is slower.
Performance degrades rapidly thereafter and
DNNs with 6 or more layers fail to execute (terminate with a ``bus error").
However, the partitioned implementation scales well to even these large DNNs.

\begin{figure}
\includegraphics[width=8cm]{partitioning.png}
\caption{Comparison of \tool code with and without partitioning. x-axis denotes the number of layers of the DNN, while y-axis denotes time in seconds for the secure protocol.}
\label{fig:partitioning}
\end{figure}

>>>>

\subsection{Secure prediction}
\noindent\textbf{Standard classifiers.}
%\label{sec:shallow}
We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
decision trees, from~\cite{shafindss} on the following data sets from the UCI
machine learning repository~\cite{uci}:
 the Wisconsin Breast Cancer data set, 
Credit Approval data set, Audiology (Standardized) data set, Nursery
data set, and ECG (electrocardiogram) classification data
from~\cite{barni}.

@@

\noindent\textbf{Deep neural nets.}
We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
comparison, we consider their implementations from
\minion~\cite{minionn}, which outperforms their previous
implementations. Table~\ref{tab:nn} shows the
results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
We note that for each of these DNNs, \minion provides a
specialized protocol, while \tool uses a generic 2PC protocol
(auto) generated from high-level code.

@@

\noindent\textbf{State-of-the-art classifiers.}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
 $W$ is a $10\times 784$ matrix, and  $b$ is a $10$ length
vector.
We implement this classifier in \tool and present the results in the
first row of Table~\ref{tab:tf}. 
%We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines and cannot be evaluated using garbled circuits.

@@

\noindent\textbf{Scalability.}  To
illustrate the scalability of partitioning, we evaluate a
sequence of DNNs with and without partitioning in~\figureref{partitioning}.
All layers are identical and partitioning places each layer in a separate stage. 
For DNNs with up to 4  layers, the performance, with and without partitioning, is almost identical and the lines overlap, thereby
illustrating that partitioning does not cause any noticeable performance overheads.
Memory issues start showing up in the non-partitioned implementation of the 5 layer DNN and it is slower.
Performance degrades rapidly thereafter and
DNNs with 6 or more layers fail to execute (terminate with a ``bus error").
However, the partitioned implementation scales well to even these large DNNs.

\begin{figure}
\includegraphics[width=8cm]{partitioning.png}
\caption{Comparison of \tool code with and without partitioning. x-axis denotes the number of layers in the DNN, while y-axis denotes time in seconds for the secure protocol.}
\label{fig:partitioning}
\end{figure}

### 2018-05-09 ccs/ex.tex
para

\paragraph{Source language.}
Consider the example $w^Tx >b$ from \sectionref{intro}, where
$w$ and $b$ are the server's input (a classifier) and $x$ is
the client's input vector. \figureref{ex-sml} shows the \tool code
for this example. The code first reads the server's (resp. client's) input using
 \ls{input1} (resp. \ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
\ls{x}. Finally, the code compares \ls{acc}
with \ls{b} and outputs this only to the client using \ls{output2}.

@@

\paragraph{\tool compiler.}
\tool compiler takes as input a source program and produces a C++
program as output. \figureref{ex-aby} shows the output code for
the example in \figureref{ex-sml} -- this is also how a program written directly in
  ABY would look like. The output program
contains party-specific code for inputs and outputs
(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
computation.

@@

\paragraph{Circuit generation and evaluation.} The next step in
\tool is to
compile the output C++ code and execute it. Doing so evaluates away the
public parts of the program, including the array accesses, and
generates a  circuit comprising of arithmetic and boolean gates,
with appropriate conversion gates. The circuit is then evaluated using
a \mpc protocol.

\paragraph{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
easier for a developer to program and get the code right in
\figureref{ex-sml}, rather than the code in
\figureref{ex-aby}.
% (as we mentioned in Section~\ref{sec:intro},
%other high-level languages are limited in their use of a single,
%arithmetic or boolean, crypto back-end). 
\tool also enables the
programmer to easily modify their code, while the compiler takes care
of efficiency. For example, consider in \figureref{ex-sml} a
change from multiplication to bitwise-or in the \ls{for} loop. It
turns out that in this case, it is more efficient to do both
the addition and bitwise-or using boolean circuits (if the addition is
done using arithmetic, the conversion cost starts to take
over). In \tool, the programmer simply needs to change one operator in
the source code, and the compiler generates efficient code that uses
boolean addition. Whereas, if the programmer was writing ABY
code, she either has to sacrifice performance, or would have to
revisit many parts of the circuit and change them. In summary, \tool raises the level of abstraction for the
programmer, and generates efficient \mpc protocols automatically,
while its metatheory provides strong correctness and security
guarantees.

>>>>

\noindent\textbf{Source language.}
Consider the example $w^Tx >b$ from \sectionref{intro}, where
$w$ and $b$ are the server's input (a classifier) and $x$ is
the client's input vector. \figureref{ex-sml} shows the \tool code
for this example. The code first reads the server's (resp. client's) input using
 \ls{input1} (resp. \ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
\ls{x}. Finally, the code compares \ls{acc}
with \ls{b} and outputs this only to the client using \ls{output2}.

@@

\noindent\textbf{\tool compiler.}
\tool compiler takes as input a source program and produces a C++
program as output. \figureref{ex-aby} shows the output code for
the example in \figureref{ex-sml} -- this is also how a program written directly in
  ABY would look like. The output program
contains party-specific code for inputs and outputs
(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
computation.

@@

\noindent\textbf{Circuit generation and evaluation.} The next step in
\tool is to
compile the output C++ code and execute it. Doing so evaluates away the
public parts of the program, including the array accesses, and
generates a  circuit comprising of arithmetic and boolean gates,
with appropriate conversion gates. The circuit is then evaluated using
a \mpc protocol.

\noindent\textbf{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
easier for a developer to program and get the code right in
\figureref{ex-sml}, rather than the code in
\figureref{ex-aby}.
% (as we mentioned in Section~\ref{sec:intro},
%other high-level languages are limited in their use of a single,
%arithmetic or boolean, crypto back-end). 
\tool also enables the
programmer to easily modify their code, while the compiler takes care
of efficiency. For example, consider in \figureref{ex-sml} a
change from multiplication to bitwise-or in the \ls{for} loop. It
turns out that in this case, it is more efficient to do both
the addition and bitwise-or using boolean circuits (if the addition is
done using arithmetic, the conversion cost starts to take
over). In \tool, the programmer simply needs to change one operator in
the source code, and the compiler generates efficient code that uses
boolean addition. Whereas, if the programmer was writing ABY
code, she either has to sacrifice performance, or would have to
revisit many parts of the circuit and change them. In summary, \tool raises the level of abstraction for the
programmer, and generates efficient \mpc protocols automatically,
while its metatheory provides strong correctness and security
guarantees.

### 2018-05-09 ccs/intro.tex
half pass
para

\paragraph{Ease of programming.} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, \figureref{ex-sml}
shows an \tool source
program for $w^Tx >b$. The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with the usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, 
code optimizations can be left to the compiler,
and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\paragraph{Cryptographic-cost aware compiler.} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and generates non-cryptographic code for the array accesses.
Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an efficient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Indeed, \tool is the
first such cryptographic-cost aware compiler.

\paragraph{Scalability (secure code partitioning).} \mpc tools
often do not scale to large functionalities. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
machine memory\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data.
%\aseem{Can we
%cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we decompose the original program into a sequence of small
sub-programs, which are then sequentially processed by \tool, while
appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (i.e., the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. \tool comes
to the rescue; it automatically inserts the required instrumentation
to ensure security of these intermediate outputs (\sectionref{pipe}). As
we show in our
evaluation, partitioning allows us to program large applications
in \tool. \aseem{What about the differences from Yao pipelining? Worth
adding here that it is a backend-agnostic technique?}

@@

\paragraph{Formal guarantees.} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for all well-typed source programs, the two semantics
successfully terminate
(e.g., there are no array index out-of-bounds errors) with identical
observable outputs. For
the security theorem, we formally reduce the security
of our scheme against semi-honest (or ``honest but curious")
adversaries to the semi honest security of the  \mpc back-end. The
theorem provides protection against side-channels arising from
conditionals and memory access patterns.
 We also prove a formal security theorem against semi-honest
 adversaries for our partitioning
scheme (\sectionref{ld} and~\sectionref{pipe}).

\paragraph{Evaluation.} We have implemented \tool
using ABY~\cite{aby} as the
cryptographic back-end. We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers from Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed in the
high-level \tool source language by a non-cryptographer. Second, the
performance
of the protocols
generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
demonstrate the usefulness of partitioning by implementing an application
that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).

\paragraph{Related Work.}
Before ABY~\cite{aby}, several works have proposed
combining secure computation protocols based on homomorphic
encryption and Yao's garbled circuits
(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
and some have also developed tools that allow writing such
combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
al.~\cite{aby} observe, due
 to the high conversion cost between
homomorphic encryption and Yao's garbled circuits, these combined
protocols do not gain much performance over a single
protocol. 
Additionally, these prior works provide informal languages
or libraries that lack formal semantics and static guarantees.
Finally, we focus on the language
features necessary to implement machine learning applications.
In particular, we do not discuss declassification of
intermediate values or indexing into arrays at secret indices.
 Handling them requires additional complexity. For example,
Wysteria~\cite{wysteria} handles the former using dependent types
and ObliVM~\cite{oblivm} uses Oblivious RAM for the latter.
These situations seldomly arise in the machine learning applications that we consider:
intermediate values don't need to be declassified and arrays are traversed in an oblivious manner. We provide a detailed survey of related
work in \sectionref{related}.

>>>>

\noindent\textbf{Ease of programming.} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, \figureref{ex-sml}
shows an \tool source
program for $w^Tx >b$. The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with the usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, 
code optimizations can be left to the compiler,
and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\noindent\textbf{Cryptographic-cost aware compiler.} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and generates non-cryptographic code for the array accesses.
Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an efficient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Indeed, \tool is the
first such cryptographic-cost aware compiler.

\noindent\textbf{Scalability (secure code partitioning).} \mpc tools
often do not scale to large functionalities. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
machine memory\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data.
%\aseem{Can we
%cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we decompose the original program into a sequence of small
sub-programs, which are then sequentially processed by \tool, while
appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (i.e., the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. \tool comes
to the rescue; it automatically inserts the required instrumentation
to ensure security of these intermediate outputs (\sectionref{pipe}). As
we show in our
evaluation, partitioning allows us to program large applications
in \tool. 
%\aseem{What about the differences from Yao pipelining? Worth
%adding here that it is a backend-agnostic technique?}

@@

\noindent\textbf{Formal guarantees.} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for all well-typed source programs, the two semantics
successfully terminate
(e.g., there are no array index out-of-bounds errors) with identical
observable outputs. For
the security theorem, we formally reduce the security
of our scheme against semi-honest (or ``honest but curious")
adversaries to the semi honest security of the  \mpc back-end. The
theorem provides protection against side-channels arising from
conditionals and memory access patterns.
 We also prove a formal security theorem against semi-honest
 adversaries for our partitioning
scheme (\sectionref{ld} and~\sectionref{pipe}).

\noindent\textbf{Evaluation.} We have implemented \tool
using ABY~\cite{aby} as the
cryptographic back-end.
We compare \tool with Yao-based compilers in~\sectionref{oblivm}
and with specialized protocols in~\sectionref{eval}. 
 We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers from Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed in the
high-level \tool source language by a non-cryptographer. Second, the
performance
of the protocols
generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
demonstrate the usefulness of partitioning by implementing an application
that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).

\noindent\textbf{Related Work.}
Before ABY~\cite{aby}, several works have proposed
combining secure computation protocols based on homomorphic
encryption and Yao's garbled circuits
(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
and some have also developed tools that allow writing such
combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
al.~\cite{aby} observe, due
 to the high conversion cost between
homomorphic encryption and Yao's garbled circuits, these combined
protocols do not gain much performance over a single
protocol. 
Additionally, these prior works provide informal languages
or libraries that lack formal semantics and static guarantees.
Finally, we focus on the language
features necessary to implement machine learning applications.
In particular, we do not discuss declassification of
intermediate values or indexing into arrays at secret indices.
 Handling them requires additional complexity. For example,
Wysteria~\cite{wysteria} handles the former using dependent types
and ObliVM~\cite{oblivm} uses Oblivious RAM for the latter.
These situations seldomly arise in the machine learning applications that we consider:
intermediate values don't need to be declassified and arrays are traversed in an oblivious manner. We provide a detailed survey of related
work in \sectionref{related}.

### 2018-05-09 ccs/ld.tex
para

\paragraph{Source language.} Our language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party. We omit these for brevity.}. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

@@

\paragraph{Source semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
 Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

\paragraph{Intermediate language.} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. 
%Below we mainly focus on the bits that are different from the source
%language.
A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
$\mathcal{B}$, which denote arithmetic and boolean respectively.
Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.
Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with labels $\ell$.
Looking ahead, the label determines how the operators are
evaluated: $\mathcal{P}$-labeled operators are evaluated in-clear, $\mathcal{A}$-labeled operators generate arithmetic circuits, and
$\mathcal{B}$-labeled operators generate boolean circuits. The
form $\langle \ell \rhd m \rangle\:\widetilde{e}$ denotes coercing
$\widetilde{e}$ from label $\ell$ to label $m$.
%The statements $\widetilde{s}$ are analogous to $s$.

\paragraph{Source to intermediate compilation.} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme in our implementation.

@@

\paragraph{Evaluation to Circuits.}
\figureref{circuits} shows the syntax of circuits. A wire id range
$\wire$ denotes a set of circuit wires that carry the runtime value of a
 variable with a secret label (we will concretely define these runtime
values later as part of the circuit semantics). Circuit gates $\gate$
are wires $\wire$, input gates $\kw{in}^m_j$, multiplication gates \kw{mult}, comparison gates \kw{gt}, and multiplexer \kw{mux} gates, 
coerce gates $\langle \ell \rhd m \rangle$, and constants.
Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
gates and arrays of gates.  A circuit $\crct$ is either empty, \kw{bind}-ing of a
circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
circuits.

@@

\paragraph{Circuit semantics.} Evaluating a program in the
intermediate language produces a
circuit to be computed using a distributed \mpc protocol. With
our circuit semantics, we model the \emph{functional} aspect
of a \mpc protocol, parametrized by cryptographic encoding and
decoding functions.

@@

\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate successfully (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program. That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$. Formally, the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):

>>>>

\noindent\textbf{Source language.} Our language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party. We omit these for brevity.}. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

@@

\noindent\textbf{Source semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
 Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

\noindent\textbf{Intermediate language.} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. 
%Below we mainly focus on the bits that are different from the source
%language.
A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
$\mathcal{B}$, which denote arithmetic and boolean respectively.
Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.
Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with labels $\ell$.
Looking ahead, the label determines how the operators are
evaluated: $\mathcal{P}$-labeled operators are evaluated in-clear, $\mathcal{A}$-labeled operators generate arithmetic circuits, and
$\mathcal{B}$-labeled operators generate boolean circuits. The
form $\langle \ell \rhd m \rangle\:\widetilde{e}$ denotes coercing
$\widetilde{e}$ from label $\ell$ to label $m$.
%The statements $\widetilde{s}$ are analogous to $s$.

\noindent\textbf{Source to intermediate compilation.} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme in our implementation.

@@

\noindent\textbf{Evaluation to Circuits.}
\figureref{circuits} shows the syntax of circuits. A wire id range
$\wire$ denotes a set of circuit wires that carry the runtime value of a
 variable with a secret label (we will concretely define these runtime
values later as part of the circuit semantics). Circuit gates $\gate$
are wires $\wire$, input gates $\kw{in}^m_j$, multiplication gates \kw{mult}, comparison gates \kw{gt}, and multiplexer \kw{mux} gates, 
coerce gates $\langle \ell \rhd m \rangle$, and constants.
Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
gates and arrays of gates.  A circuit $\crct$ is either empty, \kw{bind}-ing of a
circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
circuits.

@@

\noindent\textbf{Circuit semantics.} Evaluating a program in the
intermediate language produces a
circuit to be computed using a distributed \mpc protocol. With
our circuit semantics, we model the \emph{functional} aspect
of a \mpc protocol, parametrized by cryptographic encoding and
decoding functions.

@@

\noindent\textbf{Correctness Theorem.} We prove that all well typed programs always terminate successfully (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program. That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$. Formally, the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
